---
title: "Lab 4"
#author: "Morgen Kaufman"
date: "Math 241, Week 4"
output:
  pdf_document
urlcolor: blue
---
```{r setup, include=FALSE}
# Do not modify this chunk.
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
# Put all necessary libraries here
library(tidyverse)
library(readr)
```

### Problem 1: COVID survey - interpretation

This graphic is quite difficult to interpret as there are many variables with relatively large error bars nd thus there is a lot of variably within the data. 

Some interpretations: 
1. Individuals age 30 and under tended to agree more with the statement "I trust the information that I have received about the vaccines" in comparison to those over the age of 30. 

2. Regardless of gender, individuals on average agreed with the statement "Getting the vaccine will make me feel safer at work"

3. Individuals in the medical field on average agreed the most this the statement "Based on my understanding, I believe the vaccine is safe" 

### Problem 2: COVID survey - reconstruct

```{r}
covid_survey <- read_csv("data/covid-survey.csv")
```

```{r}
covid_survey <- read_csv("data/covid-survey.csv")
covid_survey_clean <- covid_survey %>%
  filter(rowSums(!is.na(.[-1])) > 0)
dim(covid_survey_clean)
```

```{r}
colnames(covid_survey_clean) <- covid_survey_clean[1, ]
covid_survey_rmv <- covid_survey_clean[-1, ]
```

```{r}
covid_survey_relab <- covid_survey_rmv %>%
  mutate(exp_already_vax = case_when(
    exp_already_vax == 0 ~ "No",
    exp_already_vax == 1 ~ "Yes"),
    exp_flu_vax = case_when(
      exp_flu_vax == 0 ~ "No",
      exp_flu_vax == 1 ~ "Yes"),
    exp_profession = case_when(
      exp_profession == 0 ~ "Medical",
      exp_profession == 1 ~ "Nursing"),
    exp_gender = case_when(
      exp_gender == 0 ~ "Male",
      exp_gender == 1 ~ "Female",
      exp_gender == 3 ~ "Non-binary third gender",
      exp_gender == 4 ~ "Prefer not to say"),
    exp_race = case_when(
      exp_race == 1 ~ "American Indian / Alaskan Native",
      exp_race == 2 ~ "Asian",
      exp_race == 3 ~ "Black / African American",
      exp_race == 4 ~ "Native Hawaiian / Other Pacific Islander",
      exp_race == 5 ~ "White"),
    exp_ethnicity = case_when(
      exp_ethnicity == 1 ~ "Hispanic / Latino",
      exp_ethnicity == 2 ~ "Non-Hispanic/Non-Latino"),
exp_age_bin = case_when(
exp_age_bin == 0 ~ "<20",
exp_age_bin == 20 ~ "21-25",
exp_age_bin == 25 ~ "26-30",
exp_age_bin == 30 ~ ">30")
)

dim(covid_survey_relab)
```

```{r}
covid_survey_longer <- covid_survey_relab %>% 
  pivot_longer(
    cols = starts_with("exp_"), 
    names_to = "explanatory", 
    values_to = "explanatory_value"
) %>% 
  filter(!is.na(explanatory_value)) %>% 
  pivot_longer(
    cols = starts_with("resp_"), 
    names_to = "response", 
    values_to = "response_value"
)

covid_survey_longer
```

EXPLAINNNNNNNNNNNNNNNNnnnnnnnsfower8r23tu[90t24309uo]

```{r}
covid_survey_summary_stats_by_group <- covid_survey_longer %>%
group_by(explanatory, explanatory_value, response) %>%
summarise(mean = mean(as.numeric(response_value), na.rm = TRUE),
    low = quantile(as.numeric(response_value, na.rm = TRUE), 0.1, na.rm = TRUE),
    high = quantile(as.numeric(response_value, na.rm = TRUE), 0.9, na.rm = TRUE)
  )

covid_survey_summary_stats_by_group
```
```{r}
covid_survey_summary_stats_all <- covid_survey_longer %>%
group_by(response) %>%
summarise(mean = mean(as.numeric(response_value), na.rm = TRUE),
    low = quantile(as.numeric(response_value, na.rm = TRUE), 0.1, na.rm = TRUE),
    high = quantile(as.numeric(response_value, na.rm = TRUE), 0.9, na.rm = TRUE)
  )

covid_survey_summary_stats_all
```

```{r}
covid_survey_summary_stats <- bind_rows(covid_survey_summary_stats_all,covid_survey_summary_stats_by_group)
covid_survey_summary_stats
```

```{r}
covid_survey_summary_stats <- covid_survey_summary_stats %>%
  mutate( 
    response = case_when(
    response == "resp_safety" ~ "Based on my understanding, I believe the vaccine is safe",
    response == "resp_confidence_science" ~ "I am confident in the scientific vetting process for the new COVID vaccines",
    response ==   "resp_feel_safe_at_work" ~ "Getting the vaccine will make me feel safer at work",
    response == "resp_will_recommend" ~ "I will recommend the vaccine to family, friends, and community members",
    response == "resp_trust_info" ~ "I trust the information that I have received about the vaccines",
    response == "resp_concern_safety" ~ "I am concerned about the safety and side effects of the vaccine")) %>%
  mutate( 
    explanatory = case_when(
    explanatory == "exp_age_bin" ~ "Age",
    explanatory == "exp_gender" ~ "Gender",
    explanatory == "exp_race" ~ "Race",
    explanatory == "exp_ethnicity" ~ "Ethnicity",
    explanatory == "exp_profession" ~ "Profession",
    explanatory == "exp_already_vax" ~ "Had COVID vaccine",
    explanatory == "exp_flu_vax" ~ "Had flu vaccine this year"))
```

```{r, fig.width=9, fig.height=10}
covid_survey_summary_stats %>%
  na.omit() %>%
  ggplot(aes(x = mean, y = explanatory_value)) +
  geom_point()+
  geom_errorbar(aes(xmin = low, xmax = high))+
  facet_grid(rows = vars(explanatory), cols = vars(response), scales = 'free', labeller = label_wrap_gen(width = 15)) +
  labs(
    title = NULL,
    x = "Mean Likert score \n (Error bars range from 10th to 90th percentile)",
    y = NULL)
```

