---
title: "Lab 5"
#author: "Morgen Kaufman"
date: "Math 241, Week 6"
output:
  pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
# Do not modify this chunk.
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


```

```{r}
# Put all necessary libraries here
library(tidyverse)
library(rnoaa)
library(rvest)
library(httr)
library(ggplot2)
```



## Due: Friday, March 1st at 8:30am

## Goals of this lab

1. Practice grabbing data from the internet.
1. Learn to navigate new R packages.
1. Grab data from an API (either directly or using an API wrapper).
1. Scrape data from the web.


## Potential API Wrapper Packages

## Problem 1: Predicting the ~~Un~~predictable: Portland Weather

In this problem let's get comfortable with extracting data from the National Oceanic and Atmospheric Administration's (NOAA) API via the R API wrapper package `rnoaa`.

You can find more information about the datasets and variables [here](https://www.ncdc.noaa.gov/homr/reports).

```{r}
# Don't forget to install it first!
library(rnoaa)
```

a. First things first, go to [this NOAA website](https://www.ncdc.noaa.gov/cdo-web/token) to get a key emailed to you.  Then insert your key below:

```{r, eval = T}
options(noaakey = "YHYMXktJADOYAHDGHqMwoIxtVCzdnHXT")
```



b. From the National Climate Data Center (NCDC) data, use the following code to grab the stations in Multnomah County. How many stations are in Multnomah County?

```{r, eval = T}
stations <- ncdc_stations(datasetid = "GHCND", 
                          locationid = "FIPS:41051")

mult_stations <- stations$data
```

There are 25 stations in the county.

c. January was not so rainy this year, was it?  Let's grab the precipitation data for site `GHCND:US1ORMT0006` for this past January.

```{r, eval = T}
# First fill-in and run to following to determine the
# datatypeid
ncdc_datatypes(datasetid = "GHCND",
               stationid = "GHCND:US1ORMT0006")

# Now grab the data using ncdc()
precip_se_pdx <- ncdc(
                      datasetid = "GHCND",
                      stationid = "GHCND:US1ORMT0006",
                      datatypeid = "PRCP",
                      startdate = "2024-01-01",
                      enddate = "2024-01-31")

precip_se_pdx
```

d.  What is the class of `precip_se_pdx`?  Grab the data frame nested in `precip_se_pdx` and call it `precip_se_pdx_data`.
```{r}
class(precip_se_pdx)
precip_se_pdx_data <- precip_se_pdx$data

precip_se_pdx_data
```

precip_se_pdx class is listed as "ncdc_data" In the environemnt it shows up as a list 


e. Use `ymd_hms()` in the package `lubridate` to wrangle the date column into the correct format.

```{r}
library(lubridate)

precip_pdx_data_clean <- precip_se_pdx_data %>%
  mutate(date = ymd_hms(date)) 
```


f. Plot the precipitation data for this site in Portland over time.  Rumor has it that we had only one day where it didn't rain.  Is that true?

```{r}
precip_pdx_data_clean %>%
  ggplot(aes(x = day(date), y = value))+
  geom_line()
```
No this claim is false. We can see that there were multiplke days where there was no rain in January. 


g. (Bonus) Adapt the code to create a visualization that compares the precipitation data for January over the the last four years.  Do you notice any trend over time?

```{r}
precip_24 <- ncdc(
                      datasetid = "GHCND",
                      stationid = "GHCND:US1ORMT0006",
                      datatypeid = "PRCP",
                      startdate = "2024-01-01",
                      enddate = "2024-01-31")

precip_23 <- ncdc(
                      datasetid = "GHCND",
                      stationid = "GHCND:US1ORMT0006",
                      datatypeid = "PRCP",
                      startdate = "2023-01-01",
                      enddate = "2023-01-31")
precip_22 <- ncdc(
                      datasetid = "GHCND",
                      stationid = "GHCND:US1ORMT0006",
                      datatypeid = "PRCP",
                      startdate = "2022-01-01",
                      enddate = "2022-01-31")

precip_21 <- ncdc(
                      datasetid = "GHCND",
                      stationid = "GHCND:US1ORMT0006",
                      datatypeid = "PRCP",
                      startdate = "2021-01-01",
                      enddate = "2021-01-31")

precip_24_data <- precip_24$data
precip_23_data <- precip_23$data
precip_22_data <- precip_22$data
precip_21_data <- precip_21$data


precip <- bind_rows(precip_24_data, precip_23_data, precip_22_data, precip_21_data)

precip_clean <- precip %>%
  mutate(date = ymd_hms(date)) 

precip_clean %>%
  ggplot(aes(x = day(date), y = value))+
  geom_point()+
  geom_line()+
  facet_wrap(~year(date))
```

Trends: it seems as though over the last few years the amount of rain in January has evened out. It still rains white a bit in January but the highest levels in 2023 and 2024 r emuch lower than 2021 and 2022. 

## Problem 2: From API to R 

For this problem I want you to grab web data by either talking to an API directly with `httr` or using an API wrapper.  It must be an API that we have NOT used in class or in Problem 1.

Once you have grabbed the data, do any necessary wrangling to graph it and/or produce some summary statistics. Draw some conclusions from your graph and summary statistics.

### API Wrapper Suggestions for Problem 2

Here are some potential API wrapper packages.  Feel free to use one not included in this list for Problem 2.

* `gtrendsR`: "An interface for retrieving and displaying the information returned online by Google Trends is provided. Trends (number of hits) over the time as well as geographic representation of the results can be displayed."
* [`rfishbase`](https://github.com/ropensci/rfishbase): For the fish lovers
* [`darksky`](https://github.com/hrbrmstr/darksky): For global historical and current weather conditions

```{r}
library(rfishbase)
library(dplyr)
```

```{r}
spec <- fb_tbl("species")

fresh_spec <- spec %>%
  filter(Fresh == "1")

fresh_spec %>% 
  count(AnaCat) %>%
  arrange(desc(n))

fresh_spec_mig <- fresh_spec %>% 
  filter(AnaCat %in% c("non-migratory","oceanodromous", "potamodromous", "amphidromous", "anadromous", "catadromous", "oceanodromous"))

fresh_spec_mig %>%
  group_by(AnaCat) %>%
  summarise(Mean = mean(Length, na.rm = TRUE),
            Low = quantile(Length, 0.1, na.rm = TRUE),
            High = quantile(Length, 0.9, na.rm = TRUE),
            Count = n())

fresh_spec_mig %>%
  ggplot(aes(x = Length)) +
  geom_boxplot() +
  facet_wrap(~ AnaCat, scales = "free_x")
```

Conclusions: 

Migration patterns may have an effect on fish length. This would need to be tested further however as we can see nonmigratory fish tend to be much smaller than other types of fish. Additionally,  oceanodromous fish were on average the largest in length however they are also the smallest sampled group and as such this data may be skewed. Anadromous fish have the most varaibility it seems when it comes to length. 


## Problem 3: Scraping Reedie Data

Let's see what lovely data we can pull from Reed's own website.  

a. Go to [https://www.reed.edu/ir/success.html](https://www.reed.edu/ir/success.html) and scrape the two tables.

```{r}
url <- "https://www.reed.edu/ir/success.html"

tables <- url %>%
  read_html() %>%
  html_nodes(css = "table")

tbl1 <- html_table(tables[[1]], fill = TRUE)
tbl1

tbl2 <- html_table(tables[[2]], fill = TRUE)
tbl2

tbl3 <- html_table(tables[[3]], fill = TRUE)
tbl3
```



b. Grab and print out the table that is entitled "GRADUATE SCHOOLS MOST FREQUENTLY ATTENDED BY REED ALUMNI".  Why is this data frame not in a tidy format?

Not sure if I am supposed to answer this question or not but the reason its format inst tidy is because every observation doesnt have its own rows.


Each variable must have its own column. Each observation must have its own row. Each value must have its own cell.
```{r}
tbl2
```


c. Wrangle the data into a tidy format. Glimpse the resulting data frame.

```{r}
tbl2_pivot <- tbl2 %>%
   pivot_longer(c(`MBAs`, `JDs`,`PhDs`,`MDs`), names_to = "Degree",  values_to = "School")

glimpse(tbl2_pivot)
```


d. Now grab the "OCCUPATIONAL DISTRIBUTION OF ALUMNI" table and turn it into an appropriate graph.  What conclusions can we draw from the graph?

```{r}
# Hint: Use `parse_number()` within `mutate()` to fix one of the columns
  
alum <- tbl1 %>%
  mutate(X2 = parse_number(X2))

alum %>%
  ggplot(aes(x = X1, y = X2, fill = X1)) +
  geom_bar(stat = "identity") +
  labs(
    y = "Percentage", 
    x = "Occupation Category",
    title = "Occupational Distribution of Alumni") +
  theme(
    axis.text.x = element_blank(),
    legend.title = element_blank()) 
```





e. Let's now grab the Reed graduation rates over time.  Grab the data from [here](https://www.reed.edu/ir/gradrateshist.html).

Do the following to clean up the data:

* Rename the column names.  

```{r,eval = T}
url_2 <- "https://www.reed.edu/ir/gradrateshist.html"

grads <- url_2 %>%
  read_html() %>%
  html_nodes(css = "table")

grad_rates <- html_table(grads[[1]], fill = TRUE)

colnames(grad_rates) <- c("first_year_fall", "count", "four_years", "five_years", "six_years")
 
grad_parse <- grad_rates %>% mutate(
                                four_years = parse_number(four_years),
                                five_years = parse_number(five_years),
                                six_years = parse_number(six_years))
grad_parse
```


* Remove any extraneous rows.

```{r, eval = T}
grad_rm <- grad_parse %>% filter(row_number() != 1)

grad_rm
```

* Reshape the data so that there are columns for 
    + Entering class year
    + Cohort size
    + Years to graduation
    + Graduation rate

```{r}
grad_long <- grad_rm %>%
  pivot_longer(c("four_years", "five_years", "six_years"),
               names_to = "years_to_graduation",
               values_to = "graduation_rate") %>%
         mutate(cohort_size = as.numeric(count)) %>%
   select(first_year_fall, cohort_size, years_to_graduation, graduation_rate)
grad_long
```

* Make sure each column has the correct class.    


f. Create a graph comparing the graduation rates over time and draw some conclusions.

```{r}
grad_long %>%
ggplot(aes(x= first_year_fall, y= graduation_rate, group=years_to_graduation, color = years_to_graduation))+
  geom_point() +
  geom_line() +
  scale_x_discrete(breaks = seq(1982,2019, by = 5))
```

Over time, graduation rates have risen. Additionally the more years it takes for students to graduate the higher the graduation rate. 