---
title: "Lab 3"
#author: "Morgen Kaufman"
date: "Math 241, Week 3"
output:
  pdf_document
urlcolor: blue
---

```{r setup, include=FALSE}
# Do not modify this chunk.
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)


```

```{r}
libs <- c('tidyverse','knitr','viridis', 'mosaic','mosaicData','babynames', 'Lahman','nycflights13','rnoaa')
for(l in libs){
  if(!require(l,character.only = TRUE, quietly = TRUE)){
    message( sprintf('Did not have the required package << %s >> installed. Downloading now ... ',l))
    install.packages(l) 
  }
  library(l, character.only = TRUE, quietly = TRUE)
}

```



## Due: Friday, February 16th at 8:30am

## Goals of this lab

1. Practice creating functions.
1. Practice refactoring your code to make it better!  Therefore for each problem, make sure to test your functions.


### Problem 1: Subset that R Object

Here are the R objects we will use in this problem (`dats`, `pdxTreesSmall` and `ht`).    

```{r}
library(pdxTrees)
library(mosaicData)
library(dplyr)

pdxTrees <- get_pdxTrees_parks()
# Creating the objects
dats <- list(pdxTrees  = head(pdxTrees),
             Births2015 = head(Births2015),
             HELPrct = head(HELPrct), 
             sets = c("pdxTrees", "Births2015", 
                      "HELPrct"))

pdxTreesSmall  <- head(pdxTrees)

ht <- head(pdxTrees$Tree_Height, n = 15)

```

a. What are the classes of `dats`, `pdxTreesSmall` and `ht`?
dats: list
pdxTreesSmall: data frame 
ht: vector

b. Find the 10th, 11th, and 12th values of `ht`.
```{r}
ht10_12 <- ht[10:12]
print(ht10_12)
```


c. Provide the `Species` column of `pdxTrees` as a data frame with one column.
```{r}
pdxTrees_spec <- pdxTrees %>%
  select(Species)
```


d. Provide the `Species` column of `pdxTrees` as a character vector.
```{r}
spec_vector <- as.character(pdxTrees$Species)
```

e. Provide code that gives us the second entry in `sets` from `dats`.

```{r}
dats[["sets"]][2]
```

f. Subset `pdxTreesSmall` to only `Douglas-fir` and then provide the `DBH` and `Condition` of the 4th `Douglas-fir` in the dataset.  (Feel free to mix in some `tidyverse` code if you would like to.)
```{r}

DF_DBHcon4 <- pdxTreesSmall %>% 
  filter(Common_Name == "Douglas-Fir") %>%
  select("Condition", "DBH") %>% 
  slice(4)

print(DF_DBHcon4)
```




### Problem 2: Function Creation

Figure out what the following code does and then turn it into a function. For your new function, do the following:

* Test it.
* Provide default values (when appropriate).
* Use clear names for the function and arguments.
* Make sure to appropriately handle missingness.
* Generalize it by allowing the user to specify a confidence level.
* Check the inputs and stop the function if the user provides inappropriate values.


```{r}
library(pdxTrees)
thing1 <- length(pdxTrees$DBH)
thing2 <- mean(pdxTrees$DBH)
thing3 <- sd(pdxTrees$DBH)/sqrt(thing1)
thing4 <- qt(p = .975, df = thing1 - 1)
thing5 <- thing2 - thing4*thing3
thing6 <- thing2 + thing4*thing3
```

```{r}
confidence_calc <- function(data, variable) {
  n <- length(data[[variable]])
  mean <- mean(data[[variable]])
  se <- sd(data[[variable]])/sqrt(n)
  t_val <- qt(p = 0.975, df = n - 1)
  
  lower <- mean - t_val * se
  upper <- mean + t_val * se
  
  results_ci <- list(
    n = n,
    mean = mean,
    se = se,
    t_value = t_val,
    lower_ci = lower,
    upper_ci = upper
  )
  
  return(results_ci)
}

confidence_calc(pdxTrees, "DBH")
```


### Problem 3: Wrapper Function for your `ggplot`


While we (i.e. Math 241 students) all love the grammar of graphics, not everyone else does. So for this problem, we are going to practice creating wrapper functions for `ggplot2`.

Here's an example of a wrapper for a histogram.  Notice that I can't just list the variable name as an argument.  The issue has to do with how many of the `tidyverse` functions evaluate the arguments. Therefore we have to quote (`enquo()`) and then unquote (`!!`) the arguments.  (If you want to learn more, go [here](https://tidyeval.tidyverse.org/).)

```{r}
# Minimal viable product working code
ggplot(data = pdxTrees, mapping = aes(x = DBH)) +
  geom_histogram()

# Shorthand histogram function
histo <- function(data, x){
   x <- enquo(x)
  ggplot(data = data, mapping = aes(x = !!x)) +
    geom_histogram()
}

# Test it
histo(pdxTrees, DBH)
```

a. Edit `histo()` so that the user can set 

* The number of bins
* The fill color for the bars
* The color outlining the bars

```{r}
histo <- function(data, x, bins = 5, fill_color = "grey", outline_color = "black"){
   x <- enquo(x)
  ggplot(data = data, mapping = aes(x = !!x)) +
    geom_histogram(binwidth = bins, fill = fill_color, color = outline_color)
}

histo(pdxTrees, DBH, 5, "pink", "white")
```

b. Write code to create a basic scatterplot with `ggplot2`. Then write and test a function to create a basic scatterplot.

```{r}
scatter <- function(data, x, y){
   x <- enquo(x)
   y <- enquo(y)
 ggplot(data = data, mapping = aes(x = !!x, y = !!y)) +
    geom_point()
}

scatter(pdxTrees, Crown_Width_NS, Tree_Height)
```

c. Modify your scatterplot function to allow the user to ...

* Color the points by another variable.
* Set the transparency.

```{r}
scatter <- function(data, x, y, color_alt = NULL, transparency = 1.0){
   x <- enquo(x)
   y <- enquo(y)
   color_alt <- enquo(color_alt)
 ggplot(data = data, mapping = aes(x = !!x, y = !!y, color = !!color_alt)) +
    geom_point(alpha = transparency)
}

scatter(pdxTrees, Crown_Width_NS, Crown_Width_EW, Condition, 0.5)
```

d. Write and test a function for your favorite `ggplot2` graph.
```{r}
boxplot <- function(data, x, y, fill_color = "grey", transparency = 1.0){
   x <- enquo(x)
   y <- enquo(y)
 ggplot(data = data, mapping = aes(x = !!x, y = !!y)) +
    geom_boxplot(fill = fill_color, alpha = transparency)
}

boxplot(pdxTrees, Condition, Tree_Height, "pink", 0.8)
```


### Problem 4: Functioning `dplyr`


a. Take the following code and turn it into an R function to create a **conditional proportions** table.  Similar to `ggplot2`, you will need to quote and unquote the variable names.  Make sure to test your function!

```{r}
pdxTrees %>%
  count(Native, Condition) %>%
  group_by(Native) %>%
  mutate(prop = n/sum(n)) %>%
  ungroup()


prop_tbl <-  function(data, group_var, count_var) {
   group_var <- enquo(group_var)
   count_var <- enquo(count_var)
  data %>%
    count(!!group_var, !!count_var) %>%
    group_by(!!group_var) %>%
    mutate(prop = n / sum(n)) %>%
    ungroup()
}

prop_tbl(pdxTrees, Native, Condition)

```


b.  Write a function to compute the mean, median, sd, min, max, sample size, and number of missing values of a quantitative variable by the categories of another variable.  Make sure the output is a data frame (or tibble).  Don't forget to test your function.

```{r}
data_stats <- function(data, cat_var, quant_var) {
  cat_var <- enquo(cat_var)
  quant_var <- enquo(quant_var) 
  data %>%
    group_by(!!cat_var) %>%
    summarize(
      mean = mean(!!quant_var, na.rm = TRUE),
      median = median(!!quant_var,  na.rm = TRUE),
      sd = sd(!!quant_var,  na.rm = TRUE),
      min = min(!!quant_var,  na.rm = TRUE),
      max = max(!!quant_var,  na.rm = TRUE),
      count = n(),
      missing_values = sum(is.na(!!quant_var))
    ) %>%
    ungroup()
}

data_stats(pdxTrees, Condition, Crown_Width_NS)
```


### Problem 5: another `babynames` exercise

Write a function called grab_name that, when given a **name** ***and a year*** as an argument, returns the rows from the `babynames` data frame in the `babynames` package that match that name for that year (and returns an error if that name and year combination does not match any rows). Run the function once with the arguments **Ezekiel and 1883** and once with **Ezekiel and 1983**.


```{r}
grab_name <- function(myname, myyear) {
  myyear <- enquote(myyear)
  results <- babynames %>%
    filter(name == myname, year == !!myyear)
  if(nrow(results) == 0) {
    stop("No matching rows for the provided name and year combination")
  }
  return(results)
}

grab_name("Ezekiel", 1983)
```

